Evaluating machine translation for assimilation via a gap-filling task

### 2015-02-27 2015-eamt-assim/2015-eamt-assim.tex
Some intermediate corrections
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59153 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}

@@

\title{Evaluating machine translation for assimilation}

@@

As suggested by \citet{church93}, modern machine translation (MT) systems may
be divided into two broad categories according to their purpose: post-editing and assimilation systems. Both kinds may be evaluated, either to control for quality in the development process or to compare the systems. Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.

\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something?}
Despite the fact that assimilation (or gisting) is currently the most massive application of MT, few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12} These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}. As an alternative, the modification of Cloze test was introduced for assimilation evaluation, first in \citep{trosterud12} as a supplementary, and then in \citep{oregan13} as a stand-alone method. Prior to this, Cloze tests have been used to evaluate raw MT quality, see e.g. \citep{vanslype79}, \citep{somers00}. As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill the gaps in the reference (human) translation.

The gap-filling method has been successfully used to evaluate the Basque-English Apertium language pair. In this work we extend the evaluation to three more language pairs: Basque-Spanish, English-Kazakh and Tatar-Russian. For the latter two pairs, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated task creation system. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{https://github.com/Sereni/Appraise}.

We anticipate that MT systems under investigation will contribute to the users' understanding of text. We also expect to see different results depending on text domain and on the number of words left out from the task.

The paper is organised as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated. Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure. In section~\ref{sec:results} we describe and discuss the experiment results. Finally, section~\ref{sec:conclusion} draws conclusions.

\section{Methodology}
\label{sec:methodology}
\comment{FMT: This could benefit from an example i think, something like table1? }

@@

This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
of evaluating machine translation for assimilation purposes is based on the following notion: a reader's understanding 
of a given text correlates with the number of keywords they are able to correctly restore in the text. Therefore, the 
base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed. The sentence 
is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
the \emph{target language} of the machine translation system.
\comment{FMT: It would also be nice to have a screenshot here}
The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
the main sentence. There are two types of hints: first, the \emph{source}, which is the same sentence as \emph{reference}, 
also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
comes from the machine translation of the \emph{source } sentence.

\comment{EA: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?}
In the course of the experiment, the following hint combinations are offered:
\begin{description}
\item[Reference sentence only:] The participants are asked to fill the gaps not being given
any context. This task serves as a baseline score and as an indicator of phrases
that can be completed using common knowledge or language intuition (e.g.
idioms and strong collocations);
\item[Reference sentence and source sentence:] By setup, the
participants have no command of the source language, however, it may help them to fill
in proper nouns or loan words;
\item[Reference sentence and MT hint:] In addition to the task, the
participants see the source sentence translated via Apertium. This type of task is
used for measuring the contribution of machine translation to understanding the
gist of text;
\item[Reference sentence and both hints:] This task is
added for completing the possible combination of task types, and to check
for unexpected insights, should they arise.
\end{description}

@@

For each sentence, the list of candidate keywords is prepared. It constitutes all the words that 
fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
sentence length and specified gap density. All reference sentences are over 10 words in length. Finally, the required number of keywords is selected 
from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.

@@

This work has been partly funded by the Spanish Ministerio de Economía y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme. We would like to thank the volunteers who participated in the evaluations.

>>>>

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}

@@

\title{Evaluating machine translation for assimilation via a gap-filling task}

@@

As suggested by \citet{church93}, modern machine translation (MT) systems may
be divided into two broad categories according to their purpose: post-editing and assimilation systems. The output of the former is intended to be transformed into text comparable to human translation; the latter systems' goal is to enhance user's comprehension of text. Both kinds may be evaluated, either to control for quality in the development process or to compare the systems. Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.

\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something? It is from Mikel's paper, but no reference there -- should I cite Mikel?}
Despite the fact that assimilation (or gisting) is currently the most frequent application of MT, few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12}. These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}. As an alternative, the modification of Cloze \citep{taylor53} test was introduced for assimilation evaluation, first by \citep{trosterud12} as a supplementary, and then by \citep{oregan13} as a stand-alone method. Prior to this, Cloze tests have been used to evaluate raw MT quality \citep{vanslype79}, \citep{somers00}. As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation. The gap-filling task models how well users comprehend the key points of the text. Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.

\comment{EA: I understand we selected tat-rus and eng-kaz because they were in development. why did we select eu-es?}
The gap-filling method has been successfully used to evaluate the Basque-English Apertium language pair. In this work we extend the evaluation to two more language pairs: Basque-Spanish and Tatar-Russian. For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler, etc.), checks evaluators' answers, and calculates and formats generalized results. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{https://github.com/Sereni/Appraise}.

We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the evaluation will show better results in MT-assisted task type. We also expect to see different results depending on such parameters as text domain and the number of gaps in a sentence in relation to sentence length.

The paper is organised as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated. Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure. In section~\ref{sec:results} we describe and discuss the experiment results. Finally, section~\ref{sec:conclusion} draws some conclusions.

\section{Methodology}
\label{sec:methodology}

@@

\begin{table*}
  \begin{tabular}{|l|l|}
     \hline
     \textbf{10\%}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{20\%}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
     \textbf{30\%} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
     \hline
  \end{tabular}
  \caption{Example of different gap percentage settings.} 
  \label{table:percentage}
\end{table*}

This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence \comment{EA: I'm sorry, I'm not sure what else I could explain here...}
is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
the \emph{target language} of the machine translation system.
The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
the main sentence. There are two types of hints: first, the \emph{source}, which is semantically identical to \emph{reference}, 
also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.

\comment{EA: any way to fix placement of the screenshot?}
\begin{figure*}
  \centering
\includegraphics[resolution=144]{appraise-scr}
\label{figure:screenshot}
 \caption{An example set of sentences in an online environment.}
\end{figure*}

\comment{EA: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?}
In the course of the experiment, the following hint combinations are offered:
\begin{description}
\item[Reference sentence only:] The participants are asked to fill the gaps without being given
any context. This task serves as a baseline score and as an indicator of gaps
that can be completed using common knowledge or language intuition (e.g.
idioms and strong collocations);
\item[Reference sentence and source sentence:] By setup, the
participants have no command of the source language, however, it may help them to fill
in proper nouns or loan words;
\item[Reference sentence and MT hint:] In addition to the task, the
participants see the source sentence translated via the MT system Apertium \citep{forcada11}. This type of task is
used for measuring the contribution of machine translation to understanding the
gist of text;
\item[Reference sentence and both hints:] This task is
added for completing the possible combination of task types, and to check
for unexpected insights, should they arise.
\end{description}

@@

For each sentence, the list of candidate keywords is prepared. It is composed of all the words that 
fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
sentence length and specified gap density. All reference sentences are over 10 words in length. Finally, the required number of keywords is selected 
from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.

@@

This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme. We would like to thank the volunteers who participated in the evaluations.
